---
title: "calscreening2018"
author: "Group 21; Group name: 21 and me; Members: Pluto Zhang, Linfeng Hu, Cynthia Ma"
date: "2022-10-26"
output: pdf_document
---

```{r}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library(dplyr)
library(tidyr)
library(broom)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(data.table)
library(nnet)
library(VGAM)
```

```{r}
data_cal <- read_csv('calenviroscreen-3.0-results-june-2018-update.csv')
#data_cal
```


```{r}
summary(data_cal)
```

```{r}
#names(data_cal)
colnames(data_cal) <- gsub(" ", "", colnames(data_cal))
colnames(data_cal) <- gsub("\n", "", colnames(data_cal))
#names(data_cal)
```
***
columns for diseases: Asthma LowBirthWeight CardiovascularDisease

columns for socio-economic elements:
ducation LinguisticIsolation Poverty Unemployment HousingBurden Pop.Char.

columns for air pollution elements: 
Ozone PM2.5 DieselPM DrinkingWater Pesticides Tox.Release
Traffic CleanupSites GroundwaterThreats Haz.Waste Imp.WaterBodies
SolidWaste PollutionBurden
*****

```{r}
air_pollutants_vec <- names(data_cal)[12:38]
air_pollutants_vec

disease_vec <- names(data_cal)[39:44]
disease_vec

soecon_vec <- names(data_cal)[45:57]
soecon_vec
```

# EDA

## EDA for Pollution factors
```{r}
#histograms for the air pollution factors
hist(data_cal$Ozone, main='Histogram of state-level Ozone concentration in the air' )

hist(data_cal$PM2.5, main='Histogram of state-level PM 2.5 concentration in the air' )

hist(data_cal$DieselPM, main='Histogram of state-level Diesel Particle concentration in the air' )

hist(data_cal$Pesticides, main='Histogram of state-level Pesticides concentration in the air' )

hist(data_cal$Tox.Release, main='Histogram of state-level Toxin concentration in the air' )


pairs(Asthma ~ Ozone + PM2.5 + DieselPM + Pesticides + Tox.Release + Traffic+ CleanupSites + GroundwaterThreats + Haz.Waste +Imp.WaterBodies
+ SolidWaste + PollutionBurden, dat = data_cal)
```

# for other diseases

```{r}
#histograms for the other diseases factors
hist(data_cal$Asthma, main='Histogram of state-level Asthma rate(age-adjusted)' )

hist(data_cal$LowBirthWeight, main='Histogram of Low Birth Weight Prevalence' )

hist(data_cal$CardiovascularDisease, main='Histogram of state-level Cardiovascular Diseases Prevalence' )


pairs(Asthma ~ LowBirthWeight + CardiovascularDisease , dat = data_cal)
```


#for socioeconomic factors
```{r}
pairs(Asthma ~ Education + LinguisticIsolation + Poverty + Unemployment + HousingBurden, dat = data_cal)
#pairs(Pop.Char.~ Education + LinguisticIsolation + Poverty + Unemployment + HousingBurden, dat = data_cal)
```


```{r}
#forward selection
require(broom)
data_cal1 <- na.omit(data_cal)
data_cal_old <- data_cal1
#forward selection procedure using AIC values 
lm1 <- lm(Asthma ~ 1, data=data_cal1)
stepModel <- step(lm1, direction="forward",
scope=(~ Ozone + PM2.5 + DieselPM+ DrinkingWater+Pesticides+Tox.Release+Traffic +CleanupSites + GroundwaterThreats+ Haz.Waste + Imp.WaterBodies + 
SolidWaste + Education+LinguisticIsolation+Poverty+Unemployment+HousingBurden+ LowBirthWeight+CardiovascularDisease), data=data_cal1)

```
The desired model: 
Asthma ~ CardiovascularDisease + Poverty + Ozone + LowBirthWeight + 
    LinguisticIsolation + DieselPM + DrinkingWater + Unemployment + 
    GroundwaterThreats + Tox.Release + PM2.5 + CleanupSites + 
    PollutionBurden + Imp.WaterBodies + HousingBurden + Traffic + 
    Pesticides + Haz.Waste
    
```{r}
#stepwise selection
lm_stepwise <- lm(Asthma ~ Ozone + PM2.5 + DieselPM+ DrinkingWater+Pesticides+Tox.Release+Traffic +CleanupSites + GroundwaterThreats+ Haz.Waste + Imp.WaterBodies + 
SolidWaste + PollutionBurden + Education+LinguisticIsolation+Poverty+Unemployment+HousingBurden+LowBirthWeight+CardiovascularDisease, data=data_cal1)
stepModel <- step(lm_stepwise, direction="both")
```

The stepwise model selection method and forward selection method outputs match each other. Combined with our subject matter knowledge, we will proceed with modeling using the following covariates: Ozone + PM2.5 + DieselPM + DrinkingWater + Pesticides + 
    Tox.Release + Traffic + CleanupSites + GroundwaterThreats + 
    Haz.Waste + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + 
    Poverty + Unemployment + HousingBurden + LowBirthWeight + 
    CardiovascularDisease

#### Missing Data
```{r}
#check for missing data
anyNA(data_cal)
colnames(data_cal)[colSums(is.na(data_cal)) > 0]
sum(is.na(data_cal$PM2.5))
```
covariates with missing data: CES 3.0 Score, PM2.5, DrinkingWater, Traffic, LowBirthWeight, Education, LinguisticIsolation, Poverty, Unemployment, HousingBurden, Population characteristics.
```{r}
#look at rows with missing data
dat_NA <- data_cal[!complete.cases(data_cal), ]
rowSums(is.na(dat_NA))
#look at missing pattern
library(ggmice)
dat <- data_cal[,c("Asthma", "Ozone", "PM2.5", "DieselPM","Poverty", "Unemployment",  "LowBirthWeight", "LinguisticIsolation")]
plot_pattern(dat)
```
```{r}
#Attempt Multivariate Imputation
library(mice)
tempData = mice(dat, m = 5, maxit = 10, seed = 210)
summary(tempData)
#complete(tempData,action=1)
```
```{r}
lm_pureLinear = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + DrinkingWater + Pesticides + Tox.Release + Traffic + CleanupSites + GroundwaterThreats + Haz.Waste + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + Poverty + Unemployment + HousingBurden + LowBirthWeight + CardiovascularDisease, data=data_cal1)
summary(lm_pureLinear)
confint(lm_pureLinear)
#Evaluation for this simple linear model
plot(lm_pureLinear)
```
This simple linear model is unsatisfactory. We can see from the plot of Residuals vs. Fitted that there's fanning trend, no equal variance above and below the line. The QQ-plot also shows non-normality with clear deviation from the diagonal line on both ends of the fitted curve.

- pesticides, traffic, hazard waste are not statistically significant. We may consider removing these covariates.

- Housing burden as a lower significance level (higher p-value) compared to other covariates may consider removing it from linear model as well.


```{r}
#remove coefficients with lower significance level, include only if coefficients are coded "***" significant
lm_rmSig = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + DrinkingWater + Tox.Release + CleanupSites + GroundwaterThreats + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + CardiovascularDisease, data=data_cal1)
summary(lm_rmSig)
confint(lm_rmSig)
#Evaluation for this simple linear model
plot(lm_rmSig)
```

```{r}
#leave coefficients with significant level <2e-16
lm_highSig = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + DrinkingWater + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + CardiovascularDisease, data=data_cal1)
summary(lm_highSig)
confint(lm_highSig)
#Evaluation for this simple linear model
plot(lm_highSig)
```

#### Modeling
```{r}
#vector of covariates of interest
vec_cov <- c("Ozone", "PM2.5", "DieselPM", "LinguisticIsolation", "Poverty", "Unemployment", "LowBirthWeight")

#model with only covariates of interest
data_cal <- data_cal[,c("CensusTract", "TotalPopulation", "CaliforniaCounty", "ZIP", "NearbyCity(tohelpapproximatelocationonly)", "Longitude", "Latitude", "Ozone", "PM2.5", "DieselPM", "Asthma","AsthmaPctl", "LowBirthWeight","LinguisticIsolation", "Poverty", "Unemployment" )]
data_cal1 <- na.omit(data_cal)

#check proportion of missing values
sum(!complete.cases(data_cal))/nrow(data_cal)
```


##### Linear, additive, or other models (LASSO, ridge)
```{r}
library(glmnet)
library(vip)
#define outcome variable
y <- data_cal1[,"AsthmaPctl"] |> as.matrix()
#define matrix of predictor variables
x <- data_cal1[, c("Ozone", "PM2.5", "DieselPM", "LinguisticIsolation",  "Poverty", "Unemployment", "LowBirthWeight")] |> as.matrix()
elasticnet.mod = glmnet(x,y,alpha=0.5,family="gaussian")
vip(elasticnet.mod, num_features=10, geom = "point")
#now we look at the Variable importance (vip) for factors excluding Ozone
x_noOzone <- data_cal1[, c("PM2.5", "DieselPM", "LinguisticIsolation",  "Poverty", "Unemployment", "LowBirthWeight")] |> as.matrix()
elasticnet.mod.noOzone = glmnet(x_noOzone,y,alpha=0.5,family="gaussian")
vip(elasticnet.mod.noOzone, num_features=10, geom = "point")

```

```{r}
#linear model with all linear terms for 7 covariates of interest
lm_7linear = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight, data=data_cal1)
summary(lm_7linear)
plot(lm_7linear)
```

```{r}
library(splines2)
library(foreign)
library(gam)
library(Hmisc)
#add spline term to every term with significance level <2e-16 & PM2.5
model_spline=lm(AsthmaPctl ~ bSpline(Ozone,df=4) + bSpline(PM2.5, df=4) + bSpline(DieselPM, df=4) + bSpline(LinguisticIsolation, df=4) + bSpline(Poverty, df=4) + bSpline(Unemployment, df=4) + bSpline(LowBirthWeight, df=4), data=data_cal1)
summary(model_spline)
#model evaluation
plot(model_spline)
```

```{r}
#Ridge
library(dplyr)
library(MASS)
fit = lm.ridge(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight, data_cal1, lambda = seq(0, .4, 1e-3))
#view summary of model
summary(fit)
```

More flexible modeling:
```{r}
#interaction terms
lm_interAir = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*PM2.5, data=data_cal1)
lm_interDiesel = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*DieselPM, data=data_cal1)
lm_interLing = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*LinguisticIsolation, data=data_cal1)
lm_interPov = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*Poverty, data=data_cal1)
lm_interEmp = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*Unemployment, data=data_cal1)
lm_interWgt = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Ozone*LowBirthWeight, data=data_cal1)
lm_interSoc = lm(AsthmaPctl ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight + Poverty*Unemployment, data=data_cal1)

#quadratic terms
lm_quadOzone = lm(AsthmaPctl ~ Ozone + I(Ozone)^2+PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight, data=data_cal1)
lm_quadWgt = lm(AsthmaPctl ~ Ozone + I(LowBirthWeight)^2 + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight, data=data_cal1)
```

Model comparison:
```{r}
mods_linear <- list(lm_7linear, model_spline, lm_interWgt, lm_interEmp, lm_interPov, lm_interLing, lm_interDiesel, lm_interSoc, lm_interAir, lm_quadOzone, lm_quadWgt)
mod_names <- c("7Linear", "spline", "interWgt", "interEmp", "interPov", "interLing", "interDiesel", "internSoc", "interAir", "quadOzon", "quadWgt")

library(AICcmodavg)
aictab(cand.set = mods_linear, modnames = mod_names)

r_square <- c(summary(lm_7linear)$adj.r.squared, summary(model_spline)$adj.r.squared, summary(lm_interWgt)$adj.r.squared, summary(lm_interEmp)$adj.r.squared, summary(lm_interPov)$adj.r.squared, summary(lm_interLing)$adj.r.squared, summary(lm_interDiesel)$adj.r.squared, summary(lm_interSoc)$adj.r.squared, summary(lm_interAir)$adj.r.squared, summary(lm_quadOzone)$adj.r.squared, summary(lm_quadWgt)$adj.r.squared)
cbind(mod_names,r_square)

f_score <- c(summary(lm_7linear)$fstatistic[1], summary(model_spline)$fstatistic[1], summary(lm_interWgt)$fstatistic[1], summary(lm_interEmp)$fstatistic[1], summary(lm_interPov)$fstatistic[1], summary(lm_interLing)$fstatistic[1], summary(lm_interDiesel)$fstatistic[1], summary(lm_interSoc)$fstatistic[1], summary(lm_interAir)$fstatistic[1], summary(lm_quadOzone)$fstatistic[1], summary(lm_quadWgt)$fstatistic[1])
cbind(mod_names, f_score)
```
Further interpretation of lm_interAir
```{r}
summary(lm_interAir)
plot(lm_interAir)
```



##### Poisson
```{r}

library(dplyr)
library(tidyverse)
#round the Asthma rate to make sure that it's integers
dataint <- data_cal_old |> mutate(Asthma = round(Asthma)) |> mutate(logAsthma = log(Asthma))
```


```{r}
mod.poisson.full <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + DrinkingWater + Pesticides + 
    Tox.Release + Traffic + CleanupSites + GroundwaterThreats + 
    Haz.Waste + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + 
    Poverty + Unemployment + HousingBurden + LowBirthWeight + 
    CardiovascularDisease, data=dataint, family=quasipoisson)
summary(mod.poisson.full)


sort(coef(mod.poisson.full) , decreasing = TRUE)
plot(mod.poisson.full)


```
From the summary statistics of the full model, the Cardiovascular Disease, Low Birth Weight and PM 2.5 are the three major contributors to Asthma rate, using Poisson models. 
While Ozone is the factor that is negatively correlated with Asthma rate, as is shown in the linear model. 

Since pesticides, traffic, Haz.waste, Imp.WaterBodies, PollutionBurden and Housing burden variables have  relatively lower p-values they are excluded from the analysis
Since drinking water, cleanup sites and Ground Water threats have relatively very low coefficients, we exclude their effects for the next round of analysis.  

Thus, what we keep for the next round of analysis are 8 variables: PM2.5, Ozone, DieselPM, Unemployment, Poverty, Linguistic Isolation, Low Birth Weight and Cardiovascular Diseases.

checking for dispersion:

```{r}
deviance(mod.poisson.full)/mod.poisson.full$df.residual
```


Since the quotient is greater than 1, 
There exists some form of overdispersion in the data.


First, assess which variable among the remaining variables is the most influential to Asthma rate. 
```{r}

#take the logarithm of Asthma rate 

yp <- dataint[,"logAsthma"] |> as.matrix()
#define matrix of predictor variables
xp <- dataint[, c("Ozone", "DieselPM", "LinguisticIsolation",  "Poverty", "Unemployment", "LowBirthWeight", 'DrinkingWater', 'CleanupSites', 'GroundwaterThreats')] |> as.matrix()
elasticnetp = glmnet(xp,yp,alpha=0.5,family="gaussian")
vip(elasticnetp, num_features=10, geom = "point")
#now we look at the Variable importance (vip) for factors excluding Ozone
xp_noOzone <- data_cal1[, c("DieselPM", "LinguisticIsolation",  "Poverty", "Unemployment", "LowBirthWeight")] |> as.matrix()
elasticnetp.noOzone = glmnet(x_noOzone,y,alpha=0.5,family="gaussian")
vip(elasticnetp.noOzone, num_features=10, geom = "point")
```
From the vip analysis of elastic net above, Ozone is the most influential factor related to log of Asthma rate, and low birth weight is the second important to associate with Asthma rate. Thus, Ozone will be regarded as the primary variate to study with. 

Next we examine and compare the model fits:

First, we look at the model with the linear terms only:

Looking at the simple model with only the 7 linear terms of covariates and without the interaction terms:
```{r}
mod.poisson.sim <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight, data=dataint, family=poisson)
summary(mod.poisson.sim)

deviance(mod.poisson.sim)/mod.poisson.sim$df.residual
```
From the summary statistics above, all the covariates are statistically significant.

Now let's examine the interaction terms:

```{r}
#
modp.pm_ozone <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * PM2.5, data=dataint, family=poisson())
modp.ozone_diesel <-glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * DieselPM, data=dataint, family=poisson())
modp.ozone_linguisticIsolation <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * LinguisticIsolation, data=dataint, family=poisson())
modp.ozone_poverty <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * Poverty, data=dataint, family=poisson())
modp.ozone_unemployment <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * Unemployment, data=dataint, family=poisson())
modp.ozone_lowbirthweight <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * LowBirthWeight, data=dataint, family=poisson())


summary(modp.pm_ozone)$coefficients[,4][9]
summary(modp.ozone_diesel)$coefficients[,4][9]
summary(modp.ozone_linguisticIsolation)$coefficients[,4][9]
summary(modp.ozone_poverty)$coefficients[,4][9]
summary(modp.ozone_lowbirthweight)$coefficients[,4][9]



```

As can see from the table above, the interaction term of Ozone and DieselPM is at the edge of significance threshold, but all other interaction terms have p-values lower than the significance level. 
Among these interaction terms, the term of Ozone and pm2.5 has the smallest p-value. Thus, we consider it the most significant interaction term regarding Ozone. 


We try to look at the AIC values of the models:
```{r}
summary(modp.pm_ozone)$aic
summary(modp.ozone_diesel)$aic
summary(modp.ozone_linguisticIsolation)$aic
summary(modp.ozone_poverty)$aic
summary(modp.ozone_lowbirthweight)$aic
```

The model is pretty overdispersed:
```{r}
deviance(modp.pm_ozone)/modp.pm_ozone$df.residual
deviance(modp.ozone_diesel)/modp.ozone_diesel$df.residual
deviance(modp.ozone_linguisticIsolation)/modp.ozone_linguisticIsolation$df.residual
deviance(modp.ozone_poverty)/modp.ozone_poverty$df.residual
deviance(modp.ozone_lowbirthweight)/modp.ozone_lowbirthweight$df.residual
```




Next, since Ozone is the air pollution factor most likely to be directly associated with Asthma rate, we examine the effects of adding in the quadratic term. 

Incoporating the quadratic terms of Ozone into the model:

Hypothesis testing: comparing the linear simple model with the 8 variables, to the model that has the quadratic term.

Hypothesis H0: this new model is better in terms of predicting the Asthma compared to the simple model


```{r}

mod.poisson.qua <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight  + I(Ozone^2) , data=dataint, family=poisson())
summary(mod.poisson.qua)

anova(mod.poisson.sim,mod.poisson.qua, test ='Chisq')
```


Since the p-value is lower than 0.05, it's sufficient for us to reject the null hypothesis. Thus, the model with quadratic terms is better in predicting Asthma rate, as compared to the simple model. 


Now examine the adding in the interaction term of PM2.5 and Ozone into the quadratic model:
```{r}
mod.poisson.qua.pm2.5ozone <- glm(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint, family=poisson())
summary(mod.poisson.qua.pm2.5ozone)
```
```{r}
anova(mod.poisson.qua, mod.poisson.qua.pm2.5ozone, test ='Chisq')
```

from the anova tests comparing the model with the interaction term and the quadratic model without, the p-value is below the significance level. Thus, the model with the interaction term is better in predicting the Asthma rate as comapred to the model without the interaction term. 

Check for overdispersion:
```{r}
deviance(mod.poisson.qua)/mod.poisson.qua$df.residual
deviance(mod.poisson.qua.pm2.5ozone)/mod.poisson.qua.pm2.5ozone$df.residual

```
The quadratic model has smaller overdispersion quotient than the linear model, which indicates that it has better goodness of fit. 

Next: we select the best model thus far by looking at the the dispersion quotients:

Here we examine the dispersion quotients of all the models above to select for the best one:

```{r}
modnames <- c('poisson.simple','pm_ozone interaction ','ozone_diesel interaction','ozone_linguisticIsolation interaction','ozone_poverty interaction','ozone_unemployment interaction','ozone_lowbirthweight interaction','poisson quadratic ozone ','quadratic+pm2.5_ozone interaction')
poisson_model_aics <- c(summary(mod.poisson.sim)$aic,
summary(modp.pm_ozone)$aic,
summary(modp.ozone_diesel)$aic,
summary(modp.ozone_linguisticIsolation)$aic,
summary(modp.ozone_poverty)$aic,
summary(modp.ozone_unemployment)$aic,
summary(modp.ozone_lowbirthweight)$aic,
summary(mod.poisson.qua)$aic,
summary(mod.poisson.qua.pm2.5ozone)$aic)


poisson_model_dispersions <- c(
  deviance(mod.poisson.sim)/mod.poisson.sim$df.residual,
  deviance(modp.pm_ozone)/modp.pm_ozone$df.residual,
  deviance(modp.ozone_diesel)/modp.ozone_diesel$df.residual,
  deviance(modp.ozone_linguisticIsolation)/modp.ozone_linguisticIsolation$df.residual,
  deviance(modp.ozone_poverty)/modp.ozone_poverty$df.residual,
  deviance(modp.ozone_unemployment)/modp.ozone_unemployment$df.residual,
  deviance(modp.ozone_lowbirthweight)/modp.ozone_lowbirthweight$df.residual,
    deviance(mod.poisson.qua)/mod.poisson.qua$df.residual,
  deviance(mod.poisson.qua.pm2.5ozone)/mod.poisson.qua.pm2.5ozone$df.residual
    
)
poisson_vec <- cbind(modnames,poisson_model_aics,poisson_model_dispersions)
poisson_vec
```



Next: to remedy for the overdispersion effects, we consider incorporating the negative binomial model onto the most fitted model selected above. 

```{r}
modnb.poisson.sim <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight, data = dataint)
summary(modnb.poisson.sim)
```
```{r}
summary(modnb.poisson.sim)$dispersion
deviance(modnb.poisson.sim)/modnb.poisson.sim$df.residual
```
The negative binomial model has dispersion quotient of 1, and all the terms are statistically significant with p-values much smaller than the 0.05 threshold. Thus, it's a desirable model at this step. 

```{r}
modnb.pm_ozone <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * PM2.5, data=dataint)
modnb.ozone_diesel <-MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * DieselPM, data=dataint)
modnb.ozone_linguisticIsolation <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * LinguisticIsolation, data=dataint)
modnb.ozone_poverty <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * Poverty, data=dataint)
modnb.ozone_unemployment <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * Unemployment, data=dataint)
modnb.ozone_lowbirthweight <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight + Ozone * LowBirthWeight, data=dataint)


summary(modnb.pm_ozone)$coefficients[,4][9]
summary(modnb.ozone_diesel)$coefficients[,4][9]
summary(modnb.ozone_linguisticIsolation)$coefficients[,4][9]
summary(modnb.ozone_poverty)$coefficients[,4][9]
summary(modnb.ozone_lowbirthweight)$coefficients[,4][9]
```
The interaction term of ozone and PM2.5 has the smallest P-value, which is the most significant.
```{r}
summary(modnb.pm_ozone)$aic
summary(modnb.ozone_diesel)$aic
summary(modnb.ozone_linguisticIsolation)$aic
summary(modnb.ozone_poverty)$aic

summary(modnb.ozone_lowbirthweight)$aic
```


```{r}
modnb.poisson.qua <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment+ LowBirthWeight  + I(Ozone^2) , data=dataint)
summary(modnb.poisson.qua)

```
Hypothesis testing:
1. incorporating the quadratic term of ozone and compare to the nb model with only the simple linear terms 
```{r}

anova(modnb.poisson.sim,modnb.poisson.qua, test ='Chisq')
```


2. Incorporating the interaction term of ozone and PM2.5
```{r}
modnb.poisson.qua.pm2.5ozone <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint)
summary(modnb.poisson.qua.pm2.5ozone)
```

```{r}
anova(modnb.poisson.qua,modnb.poisson.qua.pm2.5ozone)
```
The P-value is significant: indicating that the interaction term is significant.

The final model: 
MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint)
    
    
```{r}
finfitted <- log(modnb.poisson.qua.pm2.5ozone$fitted.values)
finres <- modnb.poisson.qua.pm2.5ozone$residuals

x=finfitted
y = finres
plot(x=finfitted, y = finres)
#abline(h = -1, col = 'cornflowerblue', lwd = 2)
scatter.smooth(y ~ x, span = 2/3, degree = 1, xlab= "Log of Fitted values", ylab= "Residuals")
```

```{r}

plot(modnb.poisson.qua.pm2.5ozone)
library(gtsummary)
tbl1<- tbl_regression(modnb.poisson.qua.pm2.5ozone)

```

```{r}
library(broom)

tidyfit <- tidy(modnb.poisson.qua.pm2.5ozone)
tidyfit

write.csv(tidyfit, 'tidyfit.csv')
```

Make a table that incorporates all the models' original dispersion, original AIC, negative binomical dispersion and negative binomial AIC.

```{r}
nb_aics <- c(summary(modnb.poisson.sim)$aic,
summary(modnb.pm_ozone)$aic,
summary(modnb.ozone_diesel)$aic,
summary(modnb.ozone_linguisticIsolation)$aic,
summary(modnb.ozone_poverty)$aic,
summary(modnb.ozone_unemployment)$aic,
summary(modnb.ozone_lowbirthweight)$aic,
summary(modnb.poisson.qua)$aic,
summary(modnb.poisson.qua.pm2.5ozone)$aic)


nb_dispersions <- c(
  deviance(modnb.poisson.sim)/modnb.poisson.sim$df.residual,
  deviance(modnb.pm_ozone)/modnb.pm_ozone$df.residual,
  deviance(modnb.ozone_diesel)/modnb.ozone_diesel$df.residual,
  deviance(modnb.ozone_linguisticIsolation)/modnb.ozone_linguisticIsolation$df.residual,
  deviance(modnb.ozone_poverty)/modnb.ozone_poverty$df.residual,
  deviance(modnb.ozone_unemployment)/modnb.ozone_unemployment$df.residual,
  deviance(modnb.ozone_lowbirthweight)/modnb.ozone_lowbirthweight$df.residual,
  deviance(modnb.poisson.qua)/modnb.poisson.qua$df.residual,
  deviance(modnb.poisson.qua.pm2.5ozone)/modnb.poisson.qua.pm2.5ozone$df.residual
)


all_vec <- cbind(poisson_vec,nb_aics,nb_dispersions)
all_vec

```


Multinomial Regression Testing:


```{r}
#multinomial regression
data_cal <- data_cal %>% mutate(asthma_cat = case_when(
  AsthmaPctl <=25 ~ 1,
  AsthmaPctl <=50 ~ 2,
  AsthmaPctl <=75 ~ 3,
  AsthmaPctl <=100 ~ 4
))

summ.MNfit <- function(fit, digits=3){
  s <- summary(fit)
  for(i in 2:length(fit$lev)) {
##
cat("\nLevel", fit$lev[i], "vs. Level", fit$lev[1], "\n")
##
betaHat <- s$coefficients[(i-1),]
se <- s$standard.errors[(i-1),]
zStat <- betaHat / se
pval <- 2 * pnorm(abs(zStat), lower.tail=FALSE)
##
RRR <- exp(betaHat)
RRR.lower <- exp(betaHat - qnorm(0.975)*se)
RRR.upper <- exp(betaHat + qnorm(0.975)*se)
##
results <- cbind(betaHat, se, pval, RRR, RRR.lower, RRR.upper)
print(round(results, digits=digits))
}
}

mod.multi <- multinom(asthma_cat ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment + LowBirthWeight, data=data_cal)

summ.MNfit(mod.multi)

mod.ord <- vglm(asthma_cat ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment + LowBirthWeight, cumulative(parallel=TRUE, reverse=TRUE),data=data_cal)

summary(mod.ord)

mod.ord.npo <- vglm(asthma_cat ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + Poverty + Unemployment + LowBirthWeight, cumulative(parallel=FALSE, reverse=TRUE),data=data_cal)

pchisq(deviance(mod.ord)-deviance(mod.ord.npo), df=df.residual(mod.ord)-df.residual(mod.ord.npo),lower.tail=F)
#proportional odds does not hold --> use multinomial


```




#Sensitivity Analysis

The final model: 
MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint)

```{r}
#yourcodeeee
finalmodel1<- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint)

finalmodel1
#multiple imputed data
dataint_imputed <- tempData$data |> mutate(Asthma = round(Asthma)) |> mutate(logAsthma = log(Asthma))
finalmodel2 <- MASS::glm.nb(Asthma ~ Ozone + PM2.5 + DieselPM + LinguisticIsolation + 
    Poverty + Unemployment+ LowBirthWeight+ I(Ozone^2) + PM2.5*Ozone, data=dataint_imputed)

finalmodel2
```
```{r}
finalmodel1
finalmodel2
```
```{r}
deviance(finalmodel1)/finalmodel1$df.residual
deviance(finalmodel2)/finalmodel2$df.residual
```

