---
title: "calscreening2018"
author: "Group 21; Group name: 21 and me; Members: Pluto Zhang, Linfeng Hu, Cynthia Ma"
date: "2022-10-26"
output: html_document
---

```{r}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library(dplyr)
library(tidyr)
library(broom)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(data.table)
```

```{r}
data_cal <- read_csv('calenviroscreen-3.0-results-june-2018-update.csv')
data_cal
```


```{r}
summary(data_cal)
```

```{r}
#names(data_cal)
colnames(data_cal) <- gsub(" ", "", colnames(data_cal))
colnames(data_cal) <- gsub("\n", "", colnames(data_cal))
names(data_cal)
```

from the summary we can see there are at most 242 NAs 


Question: how do we remove NAs? do we need to keep NAs? 
What about 

*****
columns for diseases: Asthma LowBirthWeight CardiovascularDisease

columns for socio-economic elements:
 Education LinguisticIsolation Poverty Unemployment HousingBurden Pop.Char.

columns for air pollution elements: 
Ozone PM2.5 DieselPM DrinkingWater Pesticides Tox.Release
Traffic CleanupSites GroundwaterThreats Haz.Waste Imp.WaterBodies
SolidWaste PollutionBurden
*****

```{r}
air_pollutants_vec <- names(data_cal)[12:38]
air_pollutants_vec

disease_vec <- names(data_cal)[39:44]
disease_vec

soecon_vec <- names(data_cal)[45:57]
soecon_vec
```

# EDA

## EDA for Pollution factors
```{r}
#histograms for the air pollution factors
hist(data_cal$Ozone, main='Histogram of state-level Ozone concentration in the air' )

hist(data_cal$PM2.5, main='Histogram of state-level PM 2.5 concentration in the air' )

hist(data_cal$DieselPM, main='Histogram of state-level Diesel Particle concentration in the air' )

hist(data_cal$Pesticides, main='Histogram of state-level Pesticides concentration in the air' )

hist(data_cal$Tox.Release, main='Histogram of state-level Toxin concentration in the air' )


pairs(Asthma ~ Ozone + PM2.5 + DieselPM + Pesticides + Tox.Release , dat = data_cal)
```

# for other diseases

```{r}
#histograms for the other diseases factors
hist(data_cal$Asthma, main='Histogram of state-level Asthma rate(age-adjusted)' )

hist(data_cal$LowBirthWeight, main='Histogram of Low Birth Weight Prevalence' )

hist(data_cal$CardiovascularDisease, main='Histogram of state-level Cardiovascular Diseases Prevalence' )


pairs(Asthma ~ LowBirthWeight + CardiovascularDisease , dat = data_cal)
```


#for socioeconomic factors
```{r}
pairs(Asthma ~ Education + LinguisticIsolation + Poverty + Unemployment + HousingBurden, dat = data_cal)
#pairs(Pop.Char.~ Education + LinguisticIsolation + Poverty + Unemployment + HousingBurden, dat = data_cal)
```


```{r}
#forward selection
require(broom)
data_cal1 <- na.omit(data_cal)
#forward selection procedure using AIC values 
lm1 <- lm(Asthma ~ 1, data=data_cal1)
stepModel <- step(lm1, direction="forward",
scope=(~ Ozone + PM2.5 + DieselPM+ DrinkingWater+Pesticides+Tox.Release+Traffic +CleanupSites + GroundwaterThreats+ Haz.Waste + Imp.WaterBodies + 
SolidWaste + Education+LinguisticIsolation+Poverty+Unemployment+HousingBurden+ LowBirthWeight+CardiovascularDisease), data=data_cal1)

```
The desired model: 
Asthma ~ CardiovascularDisease + Poverty + Ozone + LowBirthWeight + 
    LinguisticIsolation + DieselPM + DrinkingWater + Unemployment + 
    GroundwaterThreats + Tox.Release + PM2.5 + CleanupSites + 
    PollutionBurden + Imp.WaterBodies + HousingBurden + Traffic + 
    Pesticides + Haz.Waste
    
```{r}
#stepwise selection
lm_stepwise <- lm(Asthma ~ Ozone + PM2.5 + DieselPM+ DrinkingWater+Pesticides+Tox.Release+Traffic +CleanupSites + GroundwaterThreats+ Haz.Waste + Imp.WaterBodies + 
SolidWaste + PollutionBurden + Education+LinguisticIsolation+Poverty+Unemployment+HousingBurden+LowBirthWeight+CardiovascularDisease, data=data_cal1)
stepModel <- step(lm_stepwise, direction="both")
```

The stepwise model selection method and forward selection method outputs match each other. Combined with our subject matter knowledge, we will proceed with modeling using the following covariates: Ozone + PM2.5 + DieselPM + DrinkingWater + Pesticides + 
    Tox.Release + Traffic + CleanupSites + GroundwaterThreats + 
    Haz.Waste + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + 
    Poverty + Unemployment + HousingBurden + LowBirthWeight + 
    CardiovascularDisease

#### Missing Data


#### Modeling
```{r}
#vector of covariates of interest
vec_cov <- c("Ozone", "PM2.5", "DieselPM", "DrinkingWater", "Pesticides", "Tox.Release", "Traffic", "CleanupSites", "GroundwaterThreats", "Haz.Waste", "Imp.WaterBodies", "PollutionBurden", "LinguisticIsolation", "Poverty", "Unemployment", "HousingBurden", "LowBirthWeight", "CardiovascularDisease")
```


##### Linear, additive, or other models (LASSO, ridge)
```{r}
lm_pureLinear = lm(Asthma ~ Ozone + PM2.5 + DieselPM + DrinkingWater + Pesticides + Tox.Release + Traffic + CleanupSites + GroundwaterThreats + Haz.Waste + Imp.WaterBodies + PollutionBurden + LinguisticIsolation + Poverty + Unemployment + HousingBurden + LowBirthWeight + CardiovascularDisease, data=data_cal1)
summary(lm_pureLinear)
confint(lm_pureLinear)
#Evaluation for this simple linear model
plot(lm_pureLinear)
plot(resid(lm_pureLinear) ~ fitted(lm_pureLinear)) # assesses linearity and equal variance
abline(h=0, col="red") # adds the zero line for reference
hist(lm_pureLinear$residuals)
```
This simple linear model is unsatisfactory. We can see from the plot of Residuals vs. Fitted that there's fanning trend, no equal variance above and below the line. The QQ-plot also shows non-normality with clear deviation from the diagonal line on both ends of the fitted curve.




##### Logistic, Multinomial, Ordinal Model
```{r}

```

##### Poisson
```{r}

```

