---
title: "calscreening2018"
author: "Pluto Zhang, Linfeng Hu, Cynthia Ma"
date: "2022-10-26"
output: html_document
---

```{r}
library(rstudioapi)
library(tidyverse)
library(gam)
library(splines)
library(splines2)  
library(dplyr)
library(tidyr)
library(broom)
library(dslabs)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(data.table)
```

```{r}
data_cal <- read_csv('calenviroscreen-3.0-results-june-2018-update.csv')
data_cal
```


```{r}
summary(data_cal)
```

```{r}
#names(data_cal)
colnames(data_cal) <- gsub(" ", "", colnames(data_cal))
colnames(data_cal) <- gsub("\n", "", colnames(data_cal))
names(data_cal)
```

from the summary we can see there are at most 242 NAs 


Question: how do we remove NAs? do we need to keep NAs? 
What about 

*****
columns for diseases: Asthma LowBirthWeight CardiovascularDisease

columns for socio economic elements:
 Education LinguisticIsolation Poverty Unemployment HousingBurden Pop.Char.

columns for air pollution elements: 
Ozone PM2.5 DieselPM DrinkingWater Pesticides Tox.Release
Traffic CleanupSites GroundwaterThreats Haz.Waste Imp.WaterBodies
SolidWaste PollutionBurden
*****

```{r}
air_pollutants_vec <- names(data_cal)[12:38]
air_pollutants_vec



disease_vec <- names(data_cal)[39:44]
disease_vec

soecon <- names(data_cal)[45:57]
soecon
```

# EDA

## EDA for Pollution factors
```{r}



#histograms for the air pollution factors
hist(data_cal$Ozone, main='Histogram of state-level Ozone concentration in the air' )

hist(data_cal$PM2.5, main='Histogram of state-level PM 2.5 concentration in the air' )

hist(data_cal$DieselPM, main='Histogram of state-level Diesel Particle concentration in the air' )

hist(data_cal$Pesticides, main='Histogram of state-level Pesticides concentration in the air' )

hist(data_cal$Tox.Release, main='Histogram of state-level Toxin concentration in the air' )


pairs(Asthma ~ Ozone + PM2.5 + DieselPM + Pesticides + Tox.Release , dat = data_cal)
```

# for other diseases

```{r}
#histograms for the air pollution factors
hist(data_cal$Ozone, main='Histogram of state-level Ozone concentration in the air' )

hist(data_cal$PM2.5, main='Histogram of state-level PM 2.5 concentration in the air' )

hist(data_cal$DieselPM, main='Histogram of state-level Diesel Particle concentration in the air' )

hist(data_cal$Pesticides, main='Histogram of state-level Pesticides concentration in the air' )

hist(data_cal$Tox.Release, main='Histogram of state-level Toxin concentration in the air' )


pairs(Asthma ~ Ozone + PM2.5 + DieselPM + Pesticides + Tox.Release , dat = data_cal)
```



```{r}
#forward selection

require(broom)
data_cal1 <- na.omit(data_cal)
#forward selection procedure using AIC values 
lm1 <- lm(Asthma ~ 1, data=data_cal1)
stepModel <- step(lm1, direction="forward",
scope=(~ Ozone + PM2.5 + DieselPM+ DrinkingWater+Pesticides+Tox.Release+Traffic +CleanupSites + GroundwaterThreats+ Haz.Waste + Imp.WaterBodies + 
SolidWaste + PollutionBurden + Education+LinguisticIsolation+Poverty+Unemployment+HousingBurden+Pop.Char.+LowBirthWeight+CardiovascularDisease), data=data_cal1)
```
The desired model: 
Asthma ~ Pop.Char. + LinguisticIsolation + DrinkingWater + CardiovascularDisease + 
    Ozone + Education + LowBirthWeight + HousingBurden + DieselPM + 
    Unemployment + Traffic + CleanupSites + Tox.Release + Imp.WaterBodies + 
    PollutionBurden + PM2.5 + GroundwaterThreats + Haz.Waste
